\ac{GP}s are a Bayesian method for regression. We consider the regression input to be real-valued scalars $x_i$ and the regression output as the value of a function $f$ at $x_i$. The complete training data will be denoted by column vectors $\mathbf{x}$ and $\mathbf{f}$. Unseen test data is denoted with $\mathbf{\hat{x}}$ and $\mathbf{\hat{f}}$.
\ac{GP}s present a non-parametric way to express prior knowledge on the space of all possible functions $f$ modeling
a regression relationship.
Formally, a GP is an infinite-dimensional extension of the multivariate Gaussian distribution.
The joint prior distribution of training and test output is defined as 
\begin{equation}
\begin{bmatrix}
\mathbf{f} \\ 
\mathbf{\hat{f}}
\end{bmatrix}
\sim \mathcal{N}\bigg(
0,
\begin{bmatrix}
K(\mathbf{x},\mathbf{x})& K(\mathbf{x},\mathbf{\hat{x}})\\ 
K(\mathbf{\hat{x}},\mathbf{x})& K(\mathbf{\hat{x}},\mathbf{\hat{x}})
\end{bmatrix}
\bigg).
\end{equation}
We sample from the predictive posterior by applying Bayes rule to the multivariate Gaussian that is determined by the \ac{GP}:
\begin{align}
\mathbf{\hat{f}} \mid \mathbf{\hat{x}},\mathbf{x},\mathbf{f} &
\sim \mathcal{N}(
K(\mathbf{\hat{x}},\mathbf{x}) K(\mathbf{x},\mathbf{x})^{-1}\mathbf{f},
K(\mathbf{\hat{x}},\mathbf{\hat{x}}) -K(\mathbf{\hat{x}},\mathbf{x}) K(\mathbf{x},\mathbf{x})^{-1}K(\mathbf{x},\mathbf{\hat{x}})\\
& = \mathcal{N}(\bm{\hat{\mu}},\mathbf{\hat{K}})
\end{align}

Often one assumes the observed regression output is noisily measured, that is, one only sees the values of $\ybf_\noisy = \mathbf{f}+ \wbf$ where $\wbf$ is Gaussian white noise with variance $\sigma_\noise^2$. This noise term can be absorbed into the covariance matrix $\mathbf{K}(\mathbf{x},\mathbf{x})$ which in the following, we will write as $\mathbf{K}$ for readability. The log-likelihood of a \ac{GP} can then be written as:
\begin{equation}
\label{eq:gplogdens}
\log P(\mathbf{f} \mid \xbf) =
-\frac12 \ybf^\top 
\mathbf{K}^{-1} \ybf
- \frac12\log \abs{\mathbf{K}}
- \frac{n}{2}\log 2\pi
\end{equation}
where $n$ is the number of data points.
Both log-likelihood and predictive posterior can be computed efficiently in a Venture SP with an algorithm that resorts to Cholesky factorization\citep[chap. 2]{rasmussen2006gaussian} where 
we compute (\ref{eq:gplogdens}) as
\begin{equation}
\log(P(\mathbf{f}\mid \mathbf{x})\coloneqq - \frac{1}{2} \mathbf{f}^\top \bm{\alpha} - \sum_i \log \mathbf{L}_{ii} - \frac{n}{2} \log 2 \pi
\end{equation}
where 
\begin{equation}
\label{eq:chol_L}
\mathbf{L} \coloneqq \text{cholesky}(\mathbf{K})
\end{equation}
and 
\begin{equation}
\label{eq:alpha}
\bm{\alpha} \coloneqq  \mathbf{L}^\top \backslash(\mathbf{L} \backslash \mathbf{f}). 
\end{equation}
This results in a computational complexity of $\mathcal{O}(n^3)$ in the number of data points for
sampling with a complexity of $n^3/6$ for (\ref{eq:chol_L}) an $n^2/2$ for (\ref{eq:alpha}). 



The covariance function (or kernel) of a \ac{GP} governs high-level properties of the observed data such as linearity, periodicity and smoothness.
It comes with few free parameters that we call hyper-parameters.
Adjusting these results in minor changes, for example with regards to when two data points are treated similar.
More drastically different covariance functions are achieved by changing the structure of the covariance function itself.
%A different type could be a linear covariance function:
%\begin{equation}
% k(x,x^\prime) = \sigma^2 (x-\ell) (x^\prime-\ell). 
%\end{equation}
Note that covariance function structures are compositional: adding or multiplying two valid covariance functions results in another valid covariance function. 







Venture includes the primitive \texttt{make\_gp}, which takes as arguments a
unary function \texttt{mean} and a binary (symmetric, positive-semidefinite)
function \texttt{cov} and produces a function $\gtt$ distributed as a Gaussian
process with the supplied mean and covariance.  For example, a function $\gtt
\sim \GP(0,\,\SE)$, where $\SE$ is a squared-exponential covariance
\[ \SE(x, x') = \sigma^2 \exp\pn{\frac{(x-x')^2}{2\ell}} \]
with $\sigma=1$ and $\ell=1$, can be instantiated as follows:
\begin{lstlisting}[language=Venture]
assume zero = make_const_func( 0.0)
assume se = make_squaredexp( 1.0, 1.0)
assume g  = make_gp( zero, se)
\end{lstlisting}
There are two ways two view $\gtt$ as a ``random function.'' In the first view,
the \texttt{assume} directive that instantiates $\gtt$ does not use any
randomness---only the subsequent calls to $\gtt$ do---and coherence constraints
are upheld by the interpreter by keeping track of which evaluations of $\gtt$
exist in the current execution trace.  Namely, if the current trace contains evaluations
of $\gtt$ at the points $x_1,\ldots,x_N$ with return values $y_1,\ldots,y_N$,
then the next evaluation of $\gtt$ (say, jointly at the points $x_{N+1}, \ldots,
x_{N+n}$) will be distributed according to the joint conditional distribution
\[
  P\pn{\big.
    \texttt(\gtt\ x_{N+1}\texttt), \ldots, \texttt(\gtt\ x_{N+n}\texttt)
    \mvert
    \texttt(\gtt\ x_i\texttt) = f_i \text{ for $i=1,\ldots,N$}}.
\]
In the second view, $\gtt$ is a randomly chosen deterministic function, chosen
from the space of all deterministic real-valued functions; in this view, the
\texttt{assume} directive contains \emph{all} the randomness, and subsequent
invocations of $\gtt$ are deterministic.  The first view is procedural and is
faithful to the computation that occurs behind the scenes in Venture.  The
second view is declarative and is faithful to notations like ``$g \sim P(g)$''
which are often used in mathematical treatments.  Because a model program could
make arbitrarily many calls to $\gtt$, and the joint distribution on the return
values of the calls could have arbitrarily high entropy, it is not
computationally possible in finite time to choose the entire function $\gtt$ all
at once as in the second view.  Thus, it stands to reason that any
computationally implementable notion of ``nonparametric random functions'' must
involve incremental random choices in one way or another, and \ac{GP}s
in Venture are no exception.
